State of the Practice for Lattice Boltzmann Method Software

### 2022-04-25 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Revise section 9

We review software repositories for the presence of a tutorial, a user manual,
documented user characteristics, and a user support model. In total 19 software
packages have a tutorial, 13 have a user manual, and 11 have both. The tutorials
vary in scope and substance, and eight include expected output. Most user
manuals are in the form of a file that can be downloaded, while some are
rendered on a webpage. Some packages (waLBerla) do not have a user manual, but
do have useful documentation distributed throughout their webpages. Five
software packages (laboetie, LIMBES, Ludwig, Musubi, Palabos) document expected
user characteristics. Users are typically scientists or engineers. Their
background is often physics, chemistry, biophysics, or mathematics. All projects
(except LIMBES) have a user support model, and many of them have multiple
avenues of user support. The most popular avenue of support is Git, followed by
email and forums. One software package (OpenLB) has an FAQ page.

@@

\subsection{Documentation as Part of the Development Process}

@@

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackSoftDevExp}:] \textbf{Lack
	of Software Development Experience:} The developer of TCLB noted a lack of
	software development experience, and others noted a need for improving
	software engineering education. Many of the team members are domain experts,
	not computer scientists or software engineers. \citet{Nguyen-HoanEtAl2010}
	noted this same trend with research software developers, with only 23\% of
	survey respondents having a computing-related background. Similarly,
	\citet{UditAndKatz2017} show that the majority (54\%) of postdocs have not
	received training in software development.  The LBM developer suggesting an
	increasing role for formal software education matches the trend observed by
	\citet{PintoEtAl2018}, where their replication of a previous study
	\citep{HannayEtAl2009}, shows a growing interest in formal training (From
	13\% of respondents in 2009 to 22\% in 2018). \citet{PintoEtAl2018} found
	that some developers feel there is a mismatch between coding skills and
	subject-matter skills.

@@

As the LBM developers illustrate, they accomplish design for change by first
identifying likely changes, either implicitly or explicitly, and second by
hiding each likely change behind a well-defined module interface.  Although it
is unclear whether the developers are aware of Parnas's work, the approach
mirrors his recommendations \citep{Parnas1972a}.

@@

Although documentation is difficult (as mentioned under pain point
\ppref{P_Documentation}), some LBM developers have realized that prioritizing
time on documentation provides a justifiable return on investment. For instance,
the interviewed developer of ESPResSo noted that all major additions to their
package had accompanying changes to artifacts and documentation. They put
considerable time into their documentation. They further commented that their
goal was to lower the entry barrier for new developers, which meant providing
considerable developer documentation. This documentation informs developers on
how to get started; orients them to the artifacts, source code, system
architecture, and build system; and, explains the coupling between the
simulation engine and the interface. In addition, ESPResSo developers included
API documentation, although they are the only one of the top five packages to do
so.

@@

\begin{itemize}
	\item Explicitly state appropriate fluid dynamics problems that the software
	uses in its models and explicitly state the software's limitations. This is
	done in the ESPResSo user guide.
	\item Provide documentation that details the background theory, or provide a
	reference to such information. This was done by several top ranked packages,
	including ESPResSo, Ludwig, LUMA, and OpenLB.
	\item Identify expected user characteristics. LIMBES, Ludwig, laboetie, and
	Palabos all did this. Explicit user characteristics are important so the
	application and documentation can be pitched toward the right user
	background \citep{SmithEtAl2007}.
	\item Keep all documentation in one location. This was done by all top
	ranked packages.
	\item Use documentation generators, like Doxygen, as done (for instance) by
	OpenLB.  Sections~\ref{SecSurfCorrectAndVerifiab} and~\ref{Sec_CompareTools}
	give more details on the packages that use generators.
	\item Provide an explanation of theoretical principles through the user interface.
	\item Add guards in the code to test user inputs for compatibility before
	proceeding with the calculations.
	\item Include documentation on FAQ (Section~\ref{Sec_CompareArtifacts}).
\end{itemize}

>>>>

We review software repositories for the presence of a tutorial, a user manual,
documented user characteristics, and a user support model. In total 19 software
packages have a tutorial, 13 have a user manual, and 11 have both. The tutorials
vary in scope and substance, and eight include expected output. Most user
manuals are in the form of a file that can be downloaded, while some are
rendered on a webpage. Some packages (waLBerla) do not have a user manual, but
do have useful documentation distributed throughout their webpages. Five
software packages (laboetie, LIMBES, Ludwig, Musubi, Palabos) document expected
user characteristics. Users are typically scientists or engineers. Their
background is often physics, chemistry, biophysics, or mathematics. All projects
(except LIMBES) have a user support model, and many of them have multiple
avenues of user support. The most popular avenue of support is Git, followed by
email and forums. One software package (OpenLB) has an FAQ (Frequently Asked
Questions) page.

@@

\subsection{Documentation as Part of the Development Process} \label{Sec_DocInDevProc}

@@

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackSoftDevExp}:] \textbf{Lack
	of Software Development Experience:} The developer of TCLB noted a lack of
	software development experience, and others noted a need for improving
	software engineering education. Many of the team members are domain experts,
	not computer scientists or software engineers. \citet{Nguyen-HoanEtAl2010}
	noted this same trend with research software developers, with only 23\% of
	survey respondents having a computing-related background. Similarly,
	\citet{UditAndKatz2017} show that the majority (54\%) of postdocs have not
	received training in software development.  The LBM developer suggesting an
	increasing role for formal software education matches the trend observed by
	\citet{PintoEtAl2018}, where their replication of a previous study
	\citep{HannayEtAl2009}, shows a growing interest in formal training (from
	13\% of respondents in 2009 to 22\% in 2018). \citet{PintoEtAl2018} found
	that some developers feel there is a mismatch between coding skills and
	subject-matter skills.

@@

As the LBM developers illustrate, they accomplish design for change by first
identifying likely changes, either implicitly or explicitly, and second by
hiding each likely change behind a well-defined module interface.  Although it
is unclear whether the developers are aware of Parnas's work, the approach
mirrors his recommendations \citep{Parnas1972a}. Section~\ref{Sec_DocInDevProc}
lists ideas for how to document the design, including the likely changes, so
that they are more visible to others.

@@

Although documentation is difficult (as mentioned under pain point
\ppref{P_Documentation}), some LBM developers have realized that prioritizing
time on documentation provides a justifiable return on investment. For instance,
the interviewed developer of ESPResSo said that all major additions to their
package had accompanying changes to artifacts and documentation. They put
considerable time into their documentation. They further commented that their
goal was to lower the entry barrier for new developers, which meant providing
considerable developer documentation. This documentation informs developers on
how to get started; orients them to the artifacts, source code, system
architecture, and build system; and, explains the coupling between the
simulation engine and the interface. In addition, ESPResSo developers included
API documentation, although they are the only one of the top five packages to do
so.

@@

\begin{itemize}
	\item Explicitly state appropriate fluid dynamics problems that the software
	uses in its models and explicitly state the software's limitations. This is
	done in the ESPResSo user guide.
	\item Provide documentation that details the background theory, or provide a
	reference to such information. This was done by several top ranked packages,
	including ESPResSo, Ludwig, LUMA, and OpenLB.
	\item Identify expected user characteristics. LIMBES, Ludwig, laboetie, and
	Palabos all did this. Explicit user characteristics are important so the
	application and documentation can be pitched toward the right user
	background \citep{SmithEtAl2007}.
	\item Keep all documentation in one location. This was done by all top
	ranked packages.
	\item Use documentation generators, like Doxygen, as done (for instance) by
	OpenLB.  Sections~\ref{SecSurfCorrectAndVerifiab} and~\ref{Sec_CompareTools}
	give more details on the packages that use generators.
	\item Provide an explanation of theoretical principles through the user
	interface.
	\item Add guards in the code to test user inputs for compatibility before
	proceeding with the calculations.
	\item Include an FAQ (Section~\ref{Sec_CompareArtifacts}).
\end{itemize}
